---
title: "Amazon road network updates"
author: "Alyson Singleton"
date: "`r Sys.Date()`"
output:
  html_document: default
  word_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
# install packages, load libraries, set themes, knit formatting
pacman::p_load(tidyverse, dplyr, kableExtra, knitr, sf, raster, terra, spData, tmap, leaflet, ggplot2, spThin, mapdata, rnaturalearth, rnaturalearthdata, brazilmaps, devtools, brazilmaps, mapview, grid, gridExtra, RColorBrewer, raster, plm, fixest, cobalt, broom, ggpubr, gridExtra, cowplot, ggspatial,pals)
theme_set(theme_bw())
knitr::opts_chunk$set(echo = F, results=F, warning=F, message=F, cache=T,
                      fig.height=5, fig.width=10)

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

library(showtext)
font_add("Arial", "/Library/Fonts/Arial.ttf")  # Use the actual file path
showtext_auto()

no_axis <- theme(axis.title=element_blank(),
                 axis.text=element_blank(),
                 axis.ticks=element_blank(),
                 panel.grid.major = element_blank())
```

```{r, load region shapefiles}
# brazil amazon from IBGE
brazil_amazon <- st_read("~/Desktop/doctorate/ch3 amazon network/data/Limites_Amazonia_Legal_2022_shp/Limites_Amazonia_Legal_2022.shp")
brazil_amazon$geometry <- st_transform(brazil_amazon$geometry, 4326)
brazil_amazon_municipalities <- st_read("~/Desktop/doctorate/ch3 amazon network/data/Mun_Amazonia_Legal_2022_shp/Mun_Amazonia_Legal_2022.shp")
brazil_amazon_municipalities$geometry <- st_transform(brazil_amazon_municipalities$geometry, 4326)
```

## Figure 1: Major highway network in the Brazilian Amazon

This figure shows the current day major highway network across the Brazilian Amazon. The data used to create this figure was compiled from OpenStreetMap (OSM) and the Global Road Inventory Project (GRIP4). I validated the roads included in this figure by cross-checking with Google Maps Satellite view. These are not yet all paved roads. These are also not all of the roads in the Brazilian Amazon, please see figures below that display regional and local road networks, but rather the set roads I propose including in my study on the impacts of highway construction on dengue transmission.

```{r, validated major highways both sources}
#remove from grip4 shapes that dont correspond to a road on gmaps
grip4_roads_brazil_amazon <- st_read("~/Desktop/doctorate/ch3 amazon network/data/GRIP4_Region2_vector_shp/grip4_roads_brazil_amazon.shp")
grip4_roads_brazil_amazon_primary <- grip4_roads_brazil_amazon[which(grip4_roads_brazil_amazon$GP_RTP %in% c('2')),]
grip4_roads_brazil_amazon_primary$id <- c(1:dim(grip4_roads_brazil_amazon_primary)[1])
shapefiles_to_remove <- c(57,220,34,31,104,57,22,102,43,188,29,37,28,27,26,174,96,112,8,9,4,10,113,63,106,260,84,1,22,146,32,40,105,36,1120,221,103,83,98,99,97)
grip4_roads_brazil_amazon_primary <- grip4_roads_brazil_amazon_primary[-shapefiles_to_remove,]
grip4_roads_brazil_amazon_primary <- grip4_roads_brazil_amazon_primary[which(grip4_roads_brazil_amazon_primary$Shape_Leng > 0.02),]

#add a few non trunk pieces that are connectors
osm_amazon_roads <- st_read("~/Desktop/doctorate/ch3 amazon network/data/osm_december/osm_amazon_roads.shp")
osm_amazon_roads$fclass <- factor(osm_amazon_roads$fclass, levels = c('trunk','primary','secondary', 'tertiary'))
#osm_amazon_roads_trunk <- osm_amazon_roads[which(osm_amazon_roads$fclass == 'trunk'),]
#osm_amazon_roads_trunk_extras <- rbind(osm_amazon_roads_trunk,osm_amazon_roads[which(osm_amazon_roads$ref %in% c('BR-155','BR-155;PA-155')),])

#frind brazilian national highways as designated by highway number (i.e., ref starts with "BR")
osm_amazon_roads_ref_br <- osm_amazon_roads[which(substr(osm_amazon_roads$ref,0,2)=="BR"),]
osm_amazon_roads_ref_br_extras <- rbind(osm_amazon_roads_ref_br,osm_amazon_roads[which(osm_amazon_roads$ref %in% c('PA-150', 'PA-150;PA-475','PA-475','PA-252;PA-475','PA-483','PA-483;PA-150')),])
osm_amazon_roads_ref_br_extras <- rbind(osm_amazon_roads_ref_br_extras,osm_amazon_roads[which(osm_amazon_roads$ref == 'PA-151' & osm_amazon_roads$fclass=='trunk'),])

grip_osm_validated_map <- ggplot() +
  geom_sf(data = brazil_amazon_municipalities, fill='white', color='lightgrey', size=.15, show.legend = FALSE) +
  geom_sf(data = brazil_amazon, fill=NA, color='black', size=.15, show.legend = FALSE) +
  geom_sf(data = grip4_roads_brazil_amazon_primary, aes(geometry = geometry, color='#CD1076FF'), linewidth=0.8, alpha=0.8, show.legend = "line") +
  geom_sf(data = osm_amazon_roads_ref_br_extras, aes(geometry = geometry, color='#FFB04FFF'), linewidth=0.8, alpha=0.8, show.legend = "line") +
  scale_color_manual(name="Source", values=c('#CD1076FF','#FFB04FFF'),
                     labels=c('GRIP4','OSM')) +
  theme_minimal() +
  no_axis +
  theme(legend.text=element_text(size=12),
        legend.title=element_text(size=14),
        legend.position='right')
grip_osm_validated_map
```

## Figure 2: Major highway projects by number

This map displays the major highway projects (n = 16) in the Amazon (for now determined qualitatively by size), colored by number/name. 

```{r, section by name/project}
#table(osm_amazon_roads_ref_br_extras$ref)
osm_amazon_roads_ref_br_extras$first_ref <- substr(as.character(osm_amazon_roads_ref_br_extras$ref), 1, 6)
osm_ref_freq <- data.frame(table(osm_amazon_roads_ref_br_extras$first_ref)); colnames(osm_ref_freq) <- c("ref","frequency")
#mapview(osm_amazon_roads_ref_br_extras)

proj_gt120 <- which(osm_amazon_roads_ref_br_extras$first_ref %in% c(as.character(osm_ref_freq$ref[which(osm_ref_freq$frequency>120)]),"BR-155"))
osm_amazon_roads_ref_br_extras_proj_gt120 <- osm_amazon_roads_ref_br_extras[proj_gt120,]

#colors for n=18
n18_colors <- c(unname(tol())[1],"blue",unname(tol())[2:5], unname(tol())[6], "gold", unname(tol())[7:11], "orange4", "orange", "red", unname(tol())[12], "pink")

proj_gt120_map <- ggplot() +
  geom_sf(data = brazil_amazon_municipalities, fill='white', color='lightgrey', size=.15, show.legend = FALSE) +
  geom_sf(data = brazil_amazon, fill=NA, color='black', size=.15, show.legend = FALSE) +
  geom_sf(data = osm_amazon_roads_ref_br_extras_proj_gt120, aes(geometry = geometry, color=first_ref), linewidth=0.8, show.legend = "line") +
#  scale_color_manual(name="OSM (fclass)", values=c('#6A359CFF','#FFB04FFF','#CD1076FF', '#6497B1FF'),
#                     labels=c('Trunk','Primary','Secondary', 'Tertiary')) +
  theme_minimal() +
  scale_colour_manual(values=n18_colors) +
  labs(color='Highway No.') +
  no_axis +
  theme(legend.text=element_text(size=12),
        legend.title=element_text(size=14),
        legend.position='right')

proj_gt120_map

#stor <- osm_amazon_roads_ref_br_extras_proj_gt130
#mapview(stor, zcol = "first_ref")

```

## Figure 3: Open Street Map roads

This plot shows all national and regional roads in the Amazon basin in the OSM dataset.

```{r, OSM maps}
#general
osm_map <- ggplot() +
  geom_sf(data = brazil_amazon_municipalities, fill='white', color='lightgrey', size=.15, show.legend = FALSE) +
  geom_sf(data = brazil_amazon, fill=NA, color='black', size=.15, show.legend = FALSE) +
  geom_sf(data = osm_amazon_roads, aes(geometry = geometry, color=fclass), linewidth=0.8, show.legend = "line") +
  scale_color_manual(name="OSM (fclass)", values=c('#6A359CFF','#FFB04FFF','#CD1076FF', '#6497B1FF'),
                     labels=c('Trunk','Primary','Secondary', 'Tertiary')) +
  theme_minimal() +
  no_axis +
  theme(legend.text=element_text(size=12),
        legend.title=element_text(size=14),
        legend.position='right')
osm_map
```

## Figure 4: Globel Road Inventory Project (GRIP4) roads

The first plot shows all roads in the Amazon basin in the GRIP4 dataset, and the second plot shows all roads that GRIP4 designated as paved as of 2011. Note that some of the roads shown below were removed from Figure 1 after the validation step (i.e., they do not actually exist).

```{r, GRIP4 maps}
grip_map_all <- ggplot() +
  geom_sf(data = brazil_amazon_municipalities, fill='white', color='lightgrey', size=.15, show.legend = FALSE) +
  geom_sf(data = brazil_amazon, fill=NA, color='black', size=.15, show.legend = FALSE) +
  geom_sf(data = grip4_roads_brazil_amazon, aes(geometry = geometry, color=GP_RTP), linewidth=0.8, show.legend = "line") +
  scale_color_manual(name="GRIP4 (fclass)", values=c('#6A359CFF','#FFB04FFF','#CD1076FF'),
                     labels=c('Primary','Secondary','Tertiary')) +
  theme_minimal() +
  no_axis +
  theme(legend.text=element_text(size=12),
        legend.title=element_text(size=14),
        legend.position='right')
panela_legend <- get_legend(grip_map_all)
grip_map_all <- grip_map_all + theme(legend.position = "none")

grip4_roads_brazil_paved_amazon <- grip4_roads_brazil_amazon[which(grip4_roads_brazil_amazon$GP_RSE == 1),]
grip4_roads_brazil_paved_amazon$GP_RTP <- as.factor(grip4_roads_brazil_paved_amazon$GP_RTP)

# #6497B1FF, #6A359CFF, #FFB04FFF, #679C35FF, #CD1076FF
grip_map_paved <- ggplot() +
  geom_sf(data = brazil_amazon_municipalities, fill='white', color='lightgrey', size=.15, show.legend = FALSE) +
  geom_sf(data = brazil_amazon, fill=NA, color='black', size=.15, show.legend = FALSE) +
  geom_sf(data = grip4_roads_brazil_paved_amazon, aes(geometry = geometry, color=GP_RTP), linewidth=0.8, show.legend = "line") +
  scale_color_manual(name="GRIP4 Paved 2011 (fclass)", values=c('#6A359CFF','#FFB04FFF','#CD1076FF'),
                     labels=c('Primary','Secondary','Tertiary')) +
  theme_minimal() +
  no_axis +
  theme(legend.text=element_text(size=12),
        legend.title=element_text(size=14),
        legend.position='right')
panelb_legend <- get_legend(grip_map_paved)
grip_map_paved <- grip_map_paved + theme(legend.position = "none")

#####################
## Fig 2all
#####################

fig2all <- grid.arrange(grip_map_all, grip_map_paved, panela_legend, panelb_legend,                     
                        ncol = 2, nrow = 2,
                        layout_matrix = rbind(c(1,2),c(3,4)), heights=c(4,1))

# fig2all <- as_ggplot(fig2all) +                                
#   draw_plot_label(label = c("A", "B"), size = 17,
#                   x = c(0.05, 0.55), y = c(0.95, 0.95)) 
# 
# fig2all
```

```{r, major highways both sources, eval=F}
#both major
grip4_roads_brazil_amazon_primary <- grip4_roads_brazil_amazon[which(grip4_roads_brazil_amazon$GP_RTP %in% c('2')),]
osm_amazon_roads_trunk <- osm_amazon_roads[which(osm_amazon_roads$fclass == 'trunk'),]

grip_osm_major_map <- ggplot() +
  geom_sf(data = brazil_amazon_municipalities, fill='white', color='lightgrey', size=.15, show.legend = FALSE) +
  geom_sf(data = brazil_amazon, fill=NA, color='black', size=.15, show.legend = FALSE) +
  geom_sf(data = grip4_roads_brazil_amazon_primary, aes(geometry = geometry, color='#CD1076FF'), linewidth=0.8, alpha=0.8, show.legend = "line") +
  geom_sf(data = osm_amazon_roads_trunk, aes(geometry = geometry, color='#FFB04FFF'), linewidth=0.8, alpha=0.8, show.legend = "line") +
  scale_color_manual(name="Source", values=c('#CD1076FF','#FFB04FFF'),
                     labels=c('GRIP4','OSM')) +
  theme_minimal() +
  no_axis +
  theme(legend.text=element_text(size=12),
        legend.title=element_text(size=14),
        legend.position='right')
grip_osm_major_map
```

